---
title: "Sewanee Climate Report"
author: "Kyle Jones and friends"
date: "2024-06-12"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Introduction

Analyzing the climate of Sewanee is very important and with this data we hope to specifically analyze:

- How minimum and maximum temperatures have changed over time
- How the average temperature has changed over time
- How snowfall has changed over time for the entire period

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message=FALSE)
library(tidyverse)
library(ggplot2)
library(dplyr)
#nums 3 and 5 Min/max temp over time whole period, and avg temp over time
sw <- read_csv("D:/kyle_datalab/datalab_2024v2/climate/weather.csv")

sw1 <- sw %>% select(DATE,TMIN,TMAX) %>% mutate(AVGTEMP=(TMIN+TMAX)/2)

sw$DATE <- as.Date(sw$DATE)

sw2 <- sw %>% mutate(dateyr=(format(DATE, "%Y")))
#3
ggplot(data=sw2,aes(x=dateyr,y=TMIN)) + geom_col()

ggplot(data=sw2,aes(x=dateyr,y=TMAX)) + geom_col()
#5
ggplot(data=sw2,aes(x=dateyr,y=AVGTEMP))+geom_col()

sw2 <- sw2 %>% mutate(avgtmax = sum(TMAX)/9556,avgtmin=mean(TMIN))



```

## How minimum temperature and maximum temperature has changed over time:
```{r}
ggplot(data=sw1,aes(x=DATE,y=TMIN),na.rm=TRUE) + geom_col()+labs(x="Date",y="Minimum Temperature",title="Minimum Temperature vs. Date")

ggplot(data=sw1,aes(x=DATE,y=TMAX),na.rm=TRUE) + geom_col()+labs(x="Date",y="Maximum Temperature",title="Maximum Temperature vs. Date")

ggplot(data=sw2,aes(x=dateyr,y=avgtmin),na.rm=TRUE)+geom_col()






```

## How average temperature has changed over time:
```{r}
ggplot(data=sw1,aes(x=DATE,y=AVGTEMP),na.rm=TRUE)+geom_col()+labs(x="Date",y="Average Temperature",title="Average Temperature vs. Date")
```

## How snowfall has changed over time
```{r}

#sw2 <- sw %>% 

#ggplot(data=sw2,aes(x=DATE,y=SNOW))+geom_col()
```


